### 2015-02-22 zoldatoff/tex/pattern discovery/lecture02.tex
Конспект лекций 2-й недели курса Pattern Discovery
Добавлены лекции 6.4 - 6.8
Уточнены формулы в разделе "Mining colossal patterns"
* Добавлены формулы из статьи «Mining Colossal Frequent Patterns by
Core Pattern Fusion»
* мелкие правки по форматированию

\begin{itemize}
\item \textbf{(absolute) support (count) of X:} Frequency or the number of occurrences of an itemset X
\item \textbf{(relative) support, s:} The fraction of transactions that contains X (i.e., the probability that a transaction contains X)
\item An itemset X is \textbf{frequent} if the support of X is no less than a $minsup$ threshold (denoted as $\sigma$)
\end{itemize}

%--
\subsection{Association Rules}
Association rules: $X \to Y (s, c)$:
\begin{itemize}
\item \textbf{Support}, s: The probability that a transaction contains $X \cup Y$
\item \textbf{Confidence}, c: The conditional probability that a transaction containing X also contains Y: 
\begin{equation*}
c = \frac{sup(X \cup Y)}{sup(X)}
\end{equation*}
\end{itemize}

>>>>

\begin{itemize}
\item \textbf{(absolute) support (count) of X:} Frequency or the number of occurrences of an itemset X
\item \textbf{(relative) support, s:} The fraction of transactions that contains X (i.e., the probability that a transaction contains X)
\item An itemset X is \textbf{frequent} if the support of X is no less than a $minsup$ threshold (denoted as $\sigma$): $sup(X) \geqslant \sigma$.
\end{itemize}

%--
\subsection{Association Rules}
Association rules: $X \to Y (s, c)$:
\begin{itemize}
\item \textbf{Support} (s): the probability that a transaction contains $X \cup Y$:
\begin{equation*}
\mathrm{sup}(X \to Y) = \mathrm{P}(X \cup Y)
\end{equation*}

\item \textbf{Confidence} (c): the conditional probability that a transaction containing X also contains Y: 
\begin{equation*}
c = \mathrm{P}(Y \mid X) = \frac{sup(X \cup Y)}{sup(X)}
\end{equation*}
\end{itemize}

### 2015-02-22 zoldatoff/tex/pattern discovery/lecture04.tex
Конспект лекций 2-й недели курса Pattern Discovery
Добавлены лекции 6.4 - 6.8
Уточнены формулы в разделе "Mining colossal patterns"
* Добавлены формулы из статьи «Mining Colossal Frequent Patterns by
Core Pattern Fusion»
* мелкие правки по форматированию

\subsection{Interestingness Measures: Lift and $\chi ^2$}

@@

Lift - measure of dependent/correlated events:
\begin{equation*}
lift(B,C) = \frac{c(B \to C)}{s(C)} = \frac{s(B \cup C)}{s(B) \times s(C)}
\end{equation*}

Lift(B, C) may tell how B and C are correlated:
\begin{itemize}
\item $Lift(B, C) = 1$: B and C are independent
\item $Lift(B, C) > 1$: positively correlated
\item $Lift(B, C) < 1$: negatively correlated
\end{itemize}

%--
\subsubsection{Interestingness Measure: $\chi ^2$}
\begin{equation*}
\chi^2=\sum\frac{(Observed-Expected)^2}{Expected}
\end{equation*}

@@

\begin{gather*}
AllConf(A, B) = \frac{s(A \cup B)}{\max\{s(A), s(B)\}} \\
Jaccard(A, B) = \frac{s(A \cup B)}{s(A) + s(B) - s(A \cup B)} \\
Cosine(A, B) = \frac{s(A \cup B)}{\sqrt{s(A) \times s(B)}} \\
Kulczynsky(A, B) = \frac{1}{2}\left(\frac{s(A \cup B)}{s(A)} + \frac{s(A \cup B)}{s(B)}\right)\\
MacConf(A, B)=\max\left\{\frac{s(A)}{s(A \cup B)}, \frac{s(B)}{s(A \cup B)}\right\}
\end{gather*}

\subsection{Imbalance Ratio}
IR (Imbalance Ratio): measure the imbalance of two itemsets A and B in rule implications:
\begin{equation*}
IR(A, B)=\frac{\abs{s(A)-s(B)}}{s(A) + s(B) - s(A \cup B)}
\end{equation*}

>>>>

\subsection{Interestingness Measures: Lift and $\mathbf{\chi ^2}$}

@@

\textbf{Lift} - measure of dependent/correlated events:
\begin{equation*}
\mathrm{lift}(B,C) = \frac{c(B \to C)}{s(C)} = \frac{s(B \cup C)}{s(B) \times s(C)}
\end{equation*}

Lift(B, C) may tell how B and C are correlated:
\begin{itemize}
\item $\mathrm{Lift}(B, C) = 1$: B and C are independent
\item $\mathrm{Lift}(B, C) > 1$: positively correlated
\item $\mathrm{Lift}(B, C) < 1$: negatively correlated
\end{itemize}

%--
\subsubsection{Interestingness Measure: $\mathbf{\chi ^2}$}
\begin{equation*}
\chi^2=\sum\frac{(Observed-Expected)^2}{Expected}
\end{equation*}

@@

\begin{gather*}
\mathrm{AllConf}(A, B) = \frac{s(A \cup B)}{\max\{s(A), s(B)\}} \\
\mathrm{Jaccard}(A, B) = \frac{s(A \cup B)}{s(A) + s(B) - s(A \cup B)} \\
\mathrm{Cosine}(A, B) = \frac{s(A \cup B)}{\sqrt{s(A) \times s(B)}} \\
\mathrm{Kulczynsky}(A, B) = \frac{1}{2}\left(\frac{s(A \cup B)}{s(A)} + \frac{s(A \cup B)}{s(B)}\right)\\
\mathrm{MacConf}(A, B)=\max\left\{\frac{s(A)}{s(A \cup B)}, \frac{s(B)}{s(A \cup B)}\right\}
\end{gather*}

\subsection{Imbalance Ratio}
IR (Imbalance Ratio): measure the imbalance of two itemsets A and B in rule implications:
\begin{equation*}
\mathrm{IR}(A, B)=\frac{\abs{s(A)-s(B)}}{s(A) + s(B) - s(A \cup B)}
\end{equation*}

### 2015-02-22 zoldatoff/tex/pattern discovery/lecture05.tex
Уточнены формулы в разделе "Mining colossal patterns"
* Добавлены формулы из статьи «Mining Colossal Frequent Patterns by
Core Pattern Fusion»
* мелкие правки по форматированию

%--
\subsection{Mining Multi-Dimensional Associations}
Rules can be single-dimensional or multi-dimensional:
\begin{itemize}
\item Single-dimentional: $buys(X, \text{<<milk>>}) \Rightarrow buys(X, \text{<<bread>>})$
\item Inter-dimension association rule: $age(X, \text{<<18-25>>}) \wedge occupation(X, \text{<<student>>}) \Rightarrow buys(X, \text{<<coke>>})$ \item Hybrid-dimension association rules: $age(X, \text{<<18-25>>}) \wedge buys(X, \text{<<popcorn>>}) \Rightarrow buys(X, \text{<<coke>>})$
\end{itemize}

@@

\textbf{$\mathbf{\delta}$-clustering}. For each pattern P, find all patterns which can be expressed by P and whose distance to P is within $\delta$ ($\delta$-cover). All patterns in the cluster can be represented by P = compressed patterns.\\

Method for efficient, direct mining of compressed frequent patterns: Xin et al., VLDB’05.\\

@@

Use \textbf{MMS (Maximal Marginal Significance)} for measuring the combined significance of a pattern set: Xin et al., Extracting Redundancy-Aware Top-K Patterns, KDD’06.

@@

\subsubsection{Robustness of Colossal Patterns}
For a frequent pattern $\alpha$, a subpattern $\beta$ is a $\tau$-core pattern of $\alpha$ if $\beta$ shares a similar support set with $\alpha$, i.e., 
\begin{equation*}
\frac{\abs{D_\alpha}}{\abs{D_\beta}} \geqslant \tau, 0 < \tau \leqslant 1, 
\end{equation*}
where $\tau$ is called the \textbf{core ratio}.\\

\textbf{$\mathbf{(d, \tau)}$-robustness}\footnote{Robustness - прочность}: a pattern $\alpha$ is $(d, \tau)$-robust if d is the maximum number of items that can be removed from $\alpha$ for the resulting pattern to remain a $\tau$-core pattern of $\alpha$. For a $(d, \tau)$-robust pattern $\alpha$, it has $\Omega (2^d)$ core patterns.\\

@@

>>>>

%--
\subsection{Mining Multi-Dimensional Associations}
Rules can be single-dimensional or multi-dimensional:
\begin{itemize}
\item Single-dimentional: 
\begin{equation*}
\mathrm{buys}(X, \text{<<milk>>}) \Rightarrow \mathrm{buys}(X, \text{<<bread>>})
\end{equation*}
\item Inter-dimension association rule: 
\begin{equation*}
\mathrm{age}(X, \text{<<18-25>>}) \wedge \mathrm{occupation}(X, \text{<<student>>}) \Rightarrow \mathrm{buys}(X, \text{<<coke>>})
\end{equation*}
\item Hybrid-dimension association rules: 
\begin{equation*}
\mathrm{age}(X, \text{<<18-25>>}) \wedge \mathrm{buys}(X, \text{<<popcorn>>}) \Rightarrow \mathrm{buys}(X, \text{<<coke>>})
\end{equation*}
\end{itemize}

@@

\textbf{$\mathbf{\delta}$-clustering}. For each pattern P, find all patterns which can be expressed by P and whose distance to P is within $\delta$ ($\delta$-cover). All patterns in the cluster can be represented by P = compressed patterns.\footnote{Method for efficient, direct mining of compressed frequent patterns: Xin et al., VLDB’05.}

@@

Use \textbf{MMS (Maximal Marginal Significance)} for measuring the combined significance of a pattern set.\footnote{Xin et al., Extracting Redundancy-Aware Top-K Patterns, KDD’06.}

@@

\subsubsection{Robustness of Colossal Patterns}
\textbf{Definition.} For a frequent pattern $\alpha$, a subpattern $\beta$ is a $\mathbf{\tau}$\textbf{-core pattern} of $\alpha$ if $\beta$ shares a similar support set with $\alpha$, i.e., 
\begin{equation*}
\frac{\abs{D_\alpha}}{\abs{D_\beta}} \geqslant \tau, 0 < \tau \leqslant 1, 
\end{equation*}
where $\tau$ is called the \textbf{core ratio}.\\

\textbf{Definition.} \textbf{$\mathbf{(d, \tau)}$-robustness}\footnote{Robustness - прочность}: a pattern $\alpha$ is $(d, \tau)$-robust if d is the maximum number of items that can be removed from $\alpha$ for the resulting pattern to remain a $\tau$-core pattern of $\alpha$:
\begin{equation*}
d = \max_{\beta}\{\abs{\alpha} - \abs{\beta} \mid \beta \subseteq \alpha\text{, and }\beta\text{ is a }\tau\text{-core pattern of }\alpha\}
\end{equation*}

For a pattern $\alpha$ let $C_\alpha$ be the set of all its core patterns for a specified $\tau$:
\begin{equation*}
C_\alpha = \{\beta \mid \beta \subseteq \alpha, \frac{\abs{D_\alpha}}{\abs{D_\beta}} \geqslant \tau\}
\end{equation*}

\textbf{Theorem.} For a $(d, \tau)$-robust pattern $\alpha$:
\begin{equation*}
\abs{C_\alpha} \geqslant 2^d
\end{equation*}

@@

\textbf{Theorem.} For two patterns $\beta_1, \beta_2 \in C_\alpha$ 
\begin{equation*}
Dist(\beta_1, \beta_2) \leqslant r(\tau)\text{, where }r(\tau)=1-\frac{1}{2/\tau-1}
\end{equation*}
